---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidy.outliers

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tidy.outliers)](https://CRAN.R-project.org/package=tidy.outliers)
[![Codecov test coverage](https://codecov.io/gh/brunocarlin/tidy.outliers/branch/master/graph/badge.svg)](https://codecov.io/gh/brunocarlin/tidy.outliers?branch=master)
[![R-CMD-check](https://github.com/brunocarlin/tidy.outliers/workflows/R-CMD-check/badge.svg)](https://github.com/brunocarlin/tidy.outliers/actions)
<!-- badges: end -->

The goal of tidy.outliers is to allow for easy usage of many outliers removal methods, I currently plan to implement at least:

* all methods in the [OutlierDetection package](https://cran.r-project.org/web/packages/OutlierDetection/index.html)
* method of the [lookout package](https://github.com/Sevvandi/lookout)


## Installation

You can not yet install the released version of tidy.outliers from [CRAN](https://CRAN.R-project.org) with:

``` r
#install.packages("tidy.outliers")
```

And the development version from GitHub with:

```{r eval=FALSE, include=TRUE}
# install.packages("devtools")
devtools::install_github("brunocarlin/tidy.outliers")
```


## Example Mutation

This is a basic example which shows you how to solve a common problem:


### Load Libraries

```{r libraries, echo=TRUE, warning=FALSE,message=FALSE}
library(recipes)
library(tidy.outliers)
library(OutlierDetection)
```


### Create recipe mutating probabilities

```{r recipe}
rec_obj <-
  recipe(mpg ~ ., data = mtcars) %>%
  step_outliers_maha(all_numeric(), -all_outcomes()) %>%
  prep(mtcars)
```


### Investigate results

```{r juice}
juice(rec_obj) %>% 
  select(.outliers_maha)
```

```{r tidy}
tidy(rec_obj,number = 1)
```
## Example filtering

### Create recipe filtering outliers

```{r recipe remove}
rec_obj2 <-
  recipe(mpg ~ ., data = mtcars) %>%
  step_outliers_maha(all_numeric(), -all_outcomes()) %>%
  step_outliers_remove(contains(r"(.outliers)")) %>% 
  prep(mtcars)
```


### Investigate results 

We can see that the mtcars dataset got reduced by our function

```{r juice remove}
juice(rec_obj2) %>% glimpse()
```
```{r}
mtcars %>% glimpse()
```

And we can get which were the outliers and their probability

```{r tidy remove}
tidy(rec_obj2,number = 2) %>% 
  arrange(aggregation_results %>% desc())
```
